A Formal Model of Checked C

### 2023-01-17 sandp23/appendix.tex

\myparagraph{Tiny-bignum}
Due to its small size and simplicity, \systemname changes for Tiny-bignum was chosen to be comprehensive. Furthermore, bignum\_to\_string() was moved to the sandbox due to a memory-unsafe use of sprintf(). Given that significant \systemname conversion efforts is attributed to understanding the source-code and finding the precise extent to which we choose to propagate the taintedness or to stop and give up to marshalling the data between regions, Tiny-bignum only required 4 hours. The evaluation was performed on Tiny-bignum's test suite consisting of 4 Test cases, each of which test the functionality to scale on big numbers subject to all of the supported unary and binary operations.

>>>>

\myparagraph{Tiny-bignum}
Due to its small size and simplicity, \systemname changes for Tiny-bignum was chosen to be comprehensive. Furthermore, bignum\_to\_string() was moved to the sandbox due to a memory-unsafe use of sprintf().

### 2023-01-17 sandp23/evaluation.tex

\subsubsection{Conversion Effort}
The last column shows the number of hours required to partition each application.
On average, it takes $\sim$ XX hours for each application. However, the exact time depends on the complexity of the code and pointers used in an application.
Although the absolute time is high, partitioning is just a one-time effort for each application.
We start by annotating functions and then iteratively fixing type-checker errors.
Most of the time (80\%) in all applications is spent just running the type-checker.
The type-checker stops at the first error without giving information about all the other variable that needs to be fixed. 
For instance, in the following code:
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
_t_Ptr<int> y = ...; 
int *z;
int *x;
|\textcolor{red}{\faRemove}| z = y;
x = z;
\end{minted}
The type-checker displays an error only at the first assignment.
However, to correctly fix it, we need to annotate both~\code{x} and~\code{y}.
If $N$ pointers need to be annotated, then in the worst case, we might have to run the type-checker $N$ times, annotating one additional pointer in every run.
We plan to fix this in our future work and make the type-checker display all the affected pointer variables instead of just the type errors.

@@

\input{figures/memoryoverhead}

\systemname is evaluated on six programs with the use of a pre-included test suite for each program. We intend to demonstrate \systemname's capability of enforcing spatial safety in all of the scenarios leading to modern memory vulnerabilities by re-introducing corrected memory bugs and then performing \systemname annotations on the relevant buggy code . However, this requires the bugs to be predicted before they are discovered, which is chronologically impossible. Consequently, we also annotate large sections of relevant bug-free code to mimic the developer's intuition on making \systemname annotations without any input on the bug locality. Consequently, each of the conversions for the six programs follows a varied approach.

\subsubsection{\textbf{ProFTPD}}
\systemname changes for ProFTPD were limited in extent and aimed at the exact changes required to encapsulate CVE-2010-4221. We mark the user input to the unsafe function "pr\_netio\_telnet\_gets()" as tainted (\_TPtr<char>) and propagate its tainted-ness to its callers and callees enclosed within its defined scope.

\myparagraph{Testbench:}
Our changes for the above function were policed by 24 unique API tests, which we use to benchmark the performance. For each test, our benchmark samples the delta between the call and return time of pr\_netio\_telnet\_gets(). This sampling is repeated 10 times and its mean value is reflected in the below table.

\subsubsection{\textbf{UFTPD}}
\systemname changes for UFTPD were aimed at sandboxing CVE-2020-14149 and CVE-2020-5204. CVE-2020-14149 was recorded as a NULL pointer dereference in the handle\_CWD() which could have led to a DoS in versions before 2.12, thereby, requiring us to sandbox this function. CVE-2020-5204 was recorded as a buffer overflow vulnerability in the handle\_PORT() due to sprintf() which also required us to sandbox this function. Although we could have chosen to only mark the faulty pointers as tainted, we intended to keep our changes more generic.

\myparagraph{Testbench:}
For evaluation, we manually write a script for 3 Tests that each trigger "quote CWD", "quote PORT", and FTP "get file" request 10 times in a loop. Following this we record an entry for each of these tests as a mean of recorded timestamps for 10 executions for \systemname and generic-c versions, following which we record the relative average latency across the three tests between both of these UFTPD versions as a percentage in the below table.

@@

\subsubsection{\textbf{Results}}
Our evaluation of \systemname on six programs yields the following results:
\begin{itemize}
  \item Run-time Overhead is proportional to the extent of 
annotated pointers and sandboxed code.
  \item Marshalling can be avoided if Taintedness of a pointer is propagated throughout across all its usages.
  \item Marshalling data might sometimes be relatively cost-effective instead of propagating the taintedness of a pointer throughout the program.
  \item Marshalling can be avoided by using \_TLIB function qualifier that allows tainted pointers to be passed as parameters to corresponding non-tainted pointer arguments of library functions that are assured by the developer to not leak tainted addresses.
  \item Conversion style and extent dictate the converted program's conversion efforts and is modest at a couple of days. 
  \item Annotating unsafe variables with tainted pointers is more cost-effective than sandboxing the entire function, given the function is assured to be control-flow safe.  
\end{itemize}
% 
%\leo{The following is extremely
%  weak. ``Most of them'', were there any that weren't? Which ones? For
%  the ones that were, mention github issues.}  The random generator,
%equipped with the conversion tool, successfully found a few minor
%errors in the clang compiler, most of them were already issues in the
%git bug reports. For example, we discovered that while the ternary
%operator is implemented in the compiler it cannot handle complex
%bounds types in the branches. The static analysis is not sophisticated
%enough to properly detect that both branches have the same type. While
%not precisely a bug, the clang compiler does not permit memory for
%null terminated arrays to be allocated with calloc. Although calloc
%fills all spaces in memory with null, the compiler does not recognize
%this and claims that it is an unsafe cast.

>>>>

We ensured that the partitioned programs retained their expected functionality by verifying using corresponding test suites.

\subsubsection{Conversion Effort}
The last column shows the number of hours required to partition each application.
On average, it takes $\sim$ 3.5 hours for each application. However, the exact time depends on the complexity of the code and pointers used in an application.
Although the absolute time is high, partitioning is just a one-time effort for each application.
We start by annotating functions and then iteratively fixing type-checker errors.
Most of the time (80\%) in all applications is spent just running the type-checker.
The type-checker stops at the first error without giving information about all the other variable that needs to be fixed. 
For instance, in the following code:
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
_t_Ptr<int> y = ...; 
int *z;
int *x;
|\textcolor{red}{\faRemove}| z = y;
x = z;
\end{minted}
The type-checker displays an error only at the first assignment.
However, to correctly fix it, we need to annotate both~\code{x} and~\code{y}.
If $N$ pointers need to be annotated, then in the worst case, we might have to run the type-checker $N$ times, annotating one additional pointer in every run.
We plan to fix this in our future work and make the type-checker display all the affected pointer variables instead of just the type errors.

@@

\input{figures/memoryoverhead}
\systemname is evaluated on six programs on the basis of conversion efforts and performance with the use of a pre-included test suite for each program. We intend to demonstrate \systemname's capability of enforcing spatial safety in all of the scenarios leading to modern memory vulnerabilities by re-introducing corrected memory bugs and then performing \systemname annotations on the relevant buggy code. However, this would require the bugs to be predicted before they are discovered, which is chronologically impossible. Consequently, we also annotate large sections of relevant bug-free code to mimic the developer's intuition on making \systemname annotations without any input on the bug locality.

\myparagraph{\textbf{ProFTPD}}
\systemname changes for ProFTPD were limited in extent and aimed at the exact changes required to encapsulate CVE-2010-4221. We mark the user input to the unsafe function "pr\_netio\_telnet\_gets()" as tainted (\_TPtr<char>) and propagate its tainted-ness to its callers and callees enclosed within its defined scope.

\myparagraph{Testbench:}
ProFTPD was evaluated by sampling the time taken to execute its comprehensive test-suite. The delta in run-time between \systemname ProFTPD and its corresponding unconverted version's sample was calculated as a percentage. This percentage was recorded over ten iterations and its median value has been reflected in the table \ref{table:dataset}.

\myparagraph{\textbf{UFTPD}}
\systemname changes for UFTPD were aimed at sandboxing CVE-2020-14149 and CVE-2020-5204. CVE-2020-14149 was recorded as a NULL pointer dereference in the handle\_CWD() which could have led to a DoS in versions before 2.12, thereby, requiring us to sandbox this function. CVE-2020-5204 was recorded as a buffer overflow vulnerability in the handle\_PORT() due to sprintf() which also required us to sandbox this function. Although we could have chosen to only mark the faulty pointers as tainted, we intended to keep our changes more generic.

\myparagraph{Testbench:}
Since UFTPD's test-suite lacks coverage for some of the \systemname changes, we manually write a script for 3 Tests that each trigger "quote CWD", "quote PORT", and FTP "get file" request 10 times in a loop. Following this we record an entry for each of these tests as a median of recorded timestamps for 10 executions for \systemname and generic-c versions, following which we record the relative average latency across the three tests between both of these UFTPD versions as a percentage in the table \ref{table:dataset}.

@@

\subsubsection{\textbf{\systemname Dataset Results}}
Our evaluation of \systemname on six programs yields the following results:
\begin{itemize}
  \item Run-time Overhead is proportional to the extent of 
annotated pointers and sandboxed code.
  \item Marshalling can be avoided if Taintedness of a pointer is propagated throughout across all its usages.
  \item Marshalling data might sometimes be relatively cost-effective instead of propagating the taintedness of a pointer throughout the program.
  \item Marshalling can be avoided by using \_TLIB function qualifier that allows tainted pointers to be passed as parameters to corresponding non-tainted pointer arguments of library functions that are assured by the developer to not leak tainted addresses.
  \item Conversion style and extent dictate the converted program's conversion efforts and is modest at a couple of days.   
\end{itemize}

\subsection{Security Impact}
\label{subsec:securityimpact}
The last column of the Table ~\ref{table:conversioneffort} shows the list of all spatial safety vulnerabilities in the functions that have tainted types or are isolated in the~\ucregion{}.
As expected,~\emph{all vulnerabilities} are prevented by~\systemname{}.

The symbols~\vulprevented{} and~\vulisolated{} indicate whether the vulnerability was detected by our dynamic instrumentation or isolated in the sandbox, respectively.
This shows that~\systemname{} provides an effective mechanism to prevent spatial safety vulnerabilities.

\aravind{Fix the conversion effort table by the correct symbols}

% 
%\leo{The following is extremely
%  weak. ``Most of them'', were there any that weren't? Which ones? For
%  the ones that were, mention github issues.}  The random generator,
%equipped with the conversion tool, successfully found a few minor
%errors in the clang compiler, most of them were already issues in the
%git bug reports. For example, we discovered that while the ternary
%operator is implemented in the compiler it cannot handle complex
%bounds types in the branches. The static analysis is not sophisticated
%enough to properly detect that both branches have the same type. While
%not precisely a bug, the clang compiler does not permit memory for
%null terminated arrays to be allocated with calloc. Although calloc
%fills all spaces in memory with null, the compiler does not recognize
%this and claims that it is an unsafe cast.
